### 2011-07-27 presentations/figuresandcode.tex
Templates and presentations added.

\documentclass[12pt]{article}
\usepackage{fullpage,graphicx,psfrag,url}
\usepackage[small,bf]{caption}
\usepackage{verbatim}
\setlength{\captionmargin}{30pt}

\input defs.tex
\newcommand{\sign}{\mathop{\bf sign}}

\title{Including figures and code}

\begin{document}
\maketitle

\section{Including figures}
To include a figure in your document, you can use \textbf{includegraphics}.
The example below includes a figure from figures directory.   
\begin{figure}
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/problem.eps}
\end{center}
\caption{Problem figure.}
\label{f-problem}
\end{figure}

\section{Including code}
To include code, you can use \textbf{verbatiminput}.  (This command requires
the \textbf{verbatim} package. The following command includes code from 
matlab directory. 
\verbatiminput{matlab/problem.m}

\end{document}