### 2011-07-27 presentations/group-pres_2011-08-04.tex

\documentclass[dvips,landscape]{foils}
\usepackage{graphicx,psfrag}
\input defs.tex
\raggedright
\special{! TeXDict begin /landplus90{true}store end }
\renewcommand{\oursection}[1]{
\foilhead[-1.0cm]{#1}
}

\title{Stochastic Subgradient Method}
\author{}
\MyLogo{Prof.\ S.\ Boyd, EE364b, Stanford University}
\date{}

\begin{document}
\setlength{\parskip}{0cm}
\maketitle

\BIT \itemsep -1pt
\item noisy unbiased subgradient
\item stochastic subgradient method
\item convergence proof
\item stochastic programming
\item expected value of convex function
\item on-line learning and adaptive signal processing
\EIT

\vfill

\oursection{Noisy unbiased subgradient}

\BIT
\item
random vector $\tilde g \in \reals^n$ is a
\textbf{noisy unbiased subgradient} for
$f:\reals^n \rightarrow \reals$ at $x$ if for all $z$
\[
f(z) \geq f(x)+ (\Expect \tilde g)^T(z-x)
\]
\ie, $g = \Expect \tilde g \in \partial f(x)$

\item
same as $\tilde g= g+v$, where $g \in \partial f(x)$, $\Expect v=0$

\item $v$ can represent error in
computing $g$, measurement noise, Monte Carlo sampling error, etc.
\EIT
\newpage

\BIT
\item
if $x$ is also random, $\tilde g$ is a noisy unbiased
subgradient of $f$ at $x$ if
\[
\forall z \qquad f(z) \geq f(x)+ \Expect (\tilde g|x)^T(z-x)
\]
holds almost surely

\item
same as $\Expect (\tilde g|x) \in \partial f(x)$ (a.s.)
\EIT

\oursection{Stochastic subgradient method}

\textbf{stochastic subgradient method} is
the subgradient method, using noisy unbiased subgradients

\[
x^{(k+1)} = x^{(k)} - \alpha_k \tilde g^{(k)}
\]
\BIT\itemsep -3pt
\item $x^{(k)}$ is $k$th iterate
\item $\tilde g^{(k)}$ is any noisy unbiased
subgradient of (convex) $f$ at  $x^{(k)}$, \ie,
\[
\Expect (\tilde g^{(k)}|x^{(k)})  = g^{(k)} \in \partial f(x^{(k)})
\]
\item $\alpha_k>0$ is the $k$th step size
\item define $f_\mathrm{best}^{(k)} = \min \{
f(x^{(1)}), \ldots, f(x^{(k)}) \}$
\EIT

\oursection{Assumptions}

\BIT
\item $f^\star = \inf_x f(x) > -\infty$, with $f(x^\star)=f^\star$
\item $\Expect \|g^{(k)} \|_2^2 \leq G^2$ for all $k$
\item $\Expect \|x^{(1)}-x^\star \|_2^2 \leq R^2$ (can take $=$ here)
\item step sizes are square-summable but not summable
\[
\alpha_k \geq 0, \qquad
\sum_{k=1}^\infty \alpha^2_k = \|\alpha\|_2^2< \infty,  \qquad
\sum_{k=1}^\infty \alpha_k = \infty
\]
\EIT

\vfill

these assumptions are stronger than needed, just to simplify proofs

\oursection{Convergence results}

\BIT
\item convergence in expectation:
\[
\lim _{k\rightarrow \infty} \Expect f^{(k)}_\mathrm{best} = f^\star
\]

\item convergence in probability: for any $\epsilon> 0$,
\[
\lim _{k\rightarrow \infty} \Prob (f^{(k)}_\mathrm{best} \geq f^\star+
\epsilon) = 0
\]
\item almost sure convergence:
\[
\lim _{k\rightarrow \infty} f^{(k)}_\mathrm{best} = f^\star
\]
a.s. (we won't show this)
\EIT

\oursection{Convergence proof}

\textbf{key quantity:}
\emph{expected Euclidean distance squared to the optimal set}

\vskip 3ex

$\Expect \left. \left( \|x^{(k+1)} - x^\star\|_2^2 \;\right|\; x^{(k)}
\right) = \Expect \left. \left( \|x^{(k)}-\alpha_k \tilde
g^{(k)}-x^\star\|_2^2 \;\right|\; x^{(k)} \right)$
\begin{eqnarray*}
\quad & = & \|x^{(k)}-x^\star\|_2^2
-2\alpha_k \Expect \left. \left( \tilde g^{(k)T}
(x^{(k)}-x^\star) \,\right|\, x^{(k)} \right)
        +\alpha_k^2\Expect \left. \left( \|\tilde g^{(k)} \|_2^2
\,\right|\, x^{(k)} \right) \\
&=& \|x^{(k)}-x^\star\|_2^2
        -2\alpha_k \Expect (\tilde g^{(k)} | x^{(k)})^T
(x^{(k)}-x^\star)
        +\alpha_k^2\Expect \left. \left( \|\tilde g^{(k)} \|_2^2
\;\right|\; x^{(k)} \right) \\
&\leq& \|x^{(k)}-x^\star\|_2^2
        -2\alpha_k (f(x^{(k)})-f^\star)+
\alpha_k^2 \Expect \left. \left( \|\tilde g^{(k)}\|_2^2
\;\right|\; x^{(k)} \right)
\end{eqnarray*}
using $\Expect (\tilde g^{(k)}|x^{(k)}) \in \partial f(x^{(k)})$

\newpage
now take expectation:
\[
\Expect \|x^{(k+1)}-x^\star\|_2^2
\leq \Expect \|x^{(k)}-x^\star\|_2^2
            -2\alpha_k (\Expect f(x^{(k)})-f^\star)
            +\alpha_k^2 \Expect \|\tilde g^{(k)}\|_2^2
\]

apply recursively, and use $\Expect \|\tilde g^{(k)}\|_2^2 \leq G^2$
to get
\[
\Expect \|x^{(k+1)}-x^\star\|_2^2 \leq
\Expect\|x^{(1)}-x^\star\|_2^2
-2\sum_{i=1}^k \alpha_i  (\Expect f(x^{(i)})-f^\star)
+G^2\sum_{i=1}^k \alpha_i^2
\]

%using $\Expect \|x^{(k+1)}-x^\star\|_2^2 \geq 0$,
%$\Expect \|x^{(1)}-x^\star\|_2^2 \leq R^2$,
%and $\sum_{i=1}^k \alpha_i^2 \leq \|\alpha \|_2^2$
%hence we have
%\[
%2 \sum_{i=1}^k \alpha_i  (\Expect f(x^{(i)})-f^\star)
%\leq R^2 +G^2\| \alpha \|_2^2
%\]
and so
\[
\min_{i=1,\ldots,k} (\Expect f(x^{(i)})-f^\star) \leq
\frac{R^2+G^2\|\alpha\|_2^2}{2 \sum_{i=1}^k \alpha_i}
\]

\newpage
\BIT
\item we conclude
$ \min_{i=1,\ldots,k} \Expect f(x^{(i)}) \rightarrow f^\star $

\item
Jensen's inequality and concavity of minimum yields
\[
\Expect f^{(k)}_\mathrm{best}
= \Expect \min_{i=1,\ldots,k} f(x^{(i)})
\leq \min_{i=1,\ldots,k} \Expect f(x^{(i)})
\]
so $\Expect f^{(k)}_\mathrm{best} \rightarrow f^\star$
(convergence in expectation)

\item
Markov's inequality: for $\epsilon>0$
\[
\Prob(f^{(k)}_\mathrm{best} - f^\star \geq \epsilon) \leq
\frac{\Expect (f^{(k)}_\mathrm{best} - f^\star)}{\epsilon}
\]
righthand side goes to zero, so we get convergence in probability
\EIT

\oursection{Example}
piecewise linear minimization

\[
\begin{array}{ll} \mbox{minimize} &
f(x) = \max_{i=1,\ldots,m} (a_i^T x + b_i)
\end{array}
\]

we use stochastic subgradient algorithm with noisy subgradient
\[
\tilde g^{(k)} = g^{(k)} + v^{(k)}, \qquad
g^{(k)}\in \partial f(x^{(k)})
\]
$v^{(k)}$ independent zero mean random variables

\vfill

\newpage
problem instance: $n=20$ variables, $m=100$ terms,
$f^\star \approx 1.1$, $\alpha_k=1/k$

$v^{(k)}$ are IID $\mathcal N(0,0.5I)$ ($25\%$ noise
since $ \|g \| \approx 4.5$)

\vfill
\end{document}