### 2011-10-03 presentations/fios_2011/abstract/abstract.tex
Finished figures for fios abstract submission.

\documentclass{letter}
\usepackage{graphicx}
\begin{document}
\includegraphics[width=\textwidth]{fig/fiber.jpg}
\includegraphics[width=\textwidth]{fig/mode-conv.jpg}
\includegraphics[width=\textwidth]{fig/air-core.jpg}
\end{document}