### 2011-10-14 bible/1Peter1:17-19.tex
Added questions for Bible Study.

\documentclass[12pt]{article}
\begin{document}

\section{The Father's judgement (v.~17a)}
This verse begins with, ``And if\ldots'',
    which means that it is the following truth that
    demands obedience to the command in the second half of the verse.
Peter is saying, ``if $A$ is true, then you must do $B$''.
\begin{enumerate}
\item In this verse, who is being judged, Christians or unbelievers? 
    How do we know this?
\item This verse says that God's judgement is ``according to each man's work'',
    what is meant by one's ``work''?
\item Since God judges ``without partiality'', 
    how does God's standard of judgement vary from one person to the other, 
    if at all?
\end{enumerate}

\section{Conduct yourselves in fear (v.~17b)}
\begin{enumerate}
\item As believers, what are we \emph{not} to be afraid of?
\item What are we to fear?
\end{enumerate}

\section{The precious blood of Christ (v.~18-19)}
This verse begins with ``knowing\ldots'' which means
    that Peter is giving a reason why we are to obey the command in v.~17b.
Peter is arguing, ``Do $B$ because $C$ is true''.
\begin{enumerate}
\item In what way is the life was the life we were redeemed from 
    aimless and futile?
\item In what way is the life we were redeemed to purposeful and valuable?
\item What is so precious about Christ's blood, the perfect Lamb of God?
\end{enumerate}

\end{document}