### 2011-10-16 bible/Luke12:13-48.tex
Added materials for church Bible study.

\documentclass[12pt]{article}
\begin{document}
\title{Luke 12:13-48, True Riches}
\author{Providence Bible Church Sunday School}
\maketitle
\section*{v.~15, Jesus' warning}
From v.~15, what is the relationship between 
    fullness of life and one's posessions?
\section*{v.~16-21, Parable of the rich fool}
From v.~20-21, what do we lose and what do we gain at our death?
\section*{v.~22-28, Inappropriateness of Christian worry}
From v.~28, what heart attitude produces worry over food and clothing?
\section*{v.~29-34, Seek the kingdom of God}
How does v.~33 inform us of what it means to seek the kingdom of God?
\section*{v.~29-34, Be ready}
Why is Christ's return the most important future event for the believer?
\section*{v.~41-48, God's judgement}
From v.~42-43 and v.~48, what are the two main criteria for God's
    judgement?

\section*{Supplementary: God's judgement of believers}
\begin{enumerate}
\item Based on 2 Corinthians 5:10, 
    what will God judgement toward believers be based on?
\item Based on Ephesians 2:10, how is the work of every believer determined?
\item Based on Matthew 25:14-15, where does the believer obtain the resources
    needed to accomplish his work?
\item How does Matthew 25:20-23 inform us on how God will judge believers?
\item How does 1 Corinthians 3:11-15 inform us of God's judging of believers?
\end{enumerate}
\end{document}