### 2011-11-11 bible/1Peter1:20-21.tex
Added 1 Peter 1:20-21 Bible study notes.

\documentclass[12pt]{article}
\begin{document}
\part*{1 Peter 1:20-21}
\section*{v.~18-19 (review), The high value of Jesus' blood}
\begin{enumerate}
    \item What is it that makes Jesus' blood so precious?
\end{enumerate}

\section*{v.~20, The preparation and presentation of the Lamb}
\begin{enumerate}
    \item How does the foreordained nature of Jesus' sacrifice
        exalt the preciousness of His blood?
    \item How does the withholding of Jesus' full manifestation
        exalt the preciousness of His blood?
\end{enumerate}

\section*{v.~21a, Belief in God through Jesus Christ}
This verse tells us that it was Jesus who 
    not only paid for our sins by His blood,
    but who also caused us to believe in God.
\begin{enumerate}
    \item How, specifically, does Jesus initiate faith in believers?
    \item How did this happen in your life?
\end{enumerate}

\section*{v.~21b, The result of redemption by Jesus' blood}
This verse presents the effect of Jesus' work of 
    redeeming us and causing us to believe in God, 
    which is in stark contrast to our previous aimless life (v.~18).
\begin{enumerate}
    \item What are the \emph{specific} promises of God
        that a believer puts his/her faith and hope in?
\end{enumerate}
\end{document}