### 2011-12-09 presentations/inv_dev_plan_dec-2011/paper.tex
Added theory for inverse design.

\documentclass{article}
\title{Theory for inverse design}
\author{Jesse Lu}
\begin{document}
\maketitle
\tableofcontents
\section{Introduction}
\subsection{Problem statement}
\subsection{Key insights/assumptions}
\section{Adjoint method}
\subsection{Computational cost}
\subsection{Multi-mode formulation}
\section{Alternating directions method of multipliers (ADMM)}
\subsection{Computational cost of ADMM}
\subsection{Multi-mode ADMM}
\end{document}