### 2012-01-14 bible/1Peter2:4-10.tex
Added Bible study of 1 Peter 2:4-10.

\documentclass[12pt]{article}
\begin{document}
\part*{1 Peter 2:4-10}
% Instructions:
% Answer the questions below based on
\begin{enumerate}
\section*{The analogy (4-5)}
\item How do verses 4-5 describe the following relationships?
    \begin{enumerate}
    \item Jesus and God the Father
    \item Jesus and unbelievers
    \item Jesus and believers
    \end{enumerate}
\item Chief cornerstone: In the ancient Middle-Eastern world,
    the cornerstone was the first stone laid in the building process,
    determining the position and alignment of the rest of the structure.
    \begin{enumerate}
    \item What is the spiritual house, for which Jesus is the chief cornerstone?
        See also Ephesians 2:19-22.
    \item What is the purpose of this house?
        What is the earthly house that is analogous to this spiritual house?
    \item In what ways does Jesus serve as chief cornerstone 
        to this spiritual house?
        Read Colossians 1:15-18, Ephesians 4:15-16, and Revelation 2:12-20
        as well.
    \end{enumerate}
\item How does Jesus' relationship with unbelievers determine
    how we live in the world?
\section*{The testimony of Scripture (6-8)}
\item What does it mean, as quoted in verse 6, that 
        believers will not ``be put to shame''?
    What effect does meditating on this truth have on the believer,
        as described in verse 7?
\item Evaluate the value you place on knowing and obeying Jesus Christ,
        compared to that placed on other objects and pursuits in your life.
    How must you change, in order to rightly live
        in light of what God has done for you in Jesus Christ?
\item Verse 8 describes Jesus as a source of stumbing and offense
    to unbelievers.
    \begin{enumerate}
    \item How did Jesus cause of stumbling and offense 
            for both the Jews and the Greeks of that time?
        See 1 Corinthians 1:23.
    \item In what ways is Jesus an offense and stumbling block to our culture?
    \end{enumerate}
\section*{The appointments (8-10)}
\end{enumerate}
\end{document}