### 2012-01-20 bible/1Peter1:14-2:12.tex
Added bible study.

\documentclass[12pt]{article}
\newcommand{\BE}{\begin{enumerate}}
\newcommand{\EE}{\end{enumerate}}
\newcommand{\I}{\item}
\begin{document}
\part*{1 Peter 1:14-2:12}
This study is designed to be a review of this section of Scripture,
    where Peter's focus is holy living.
Please take multiple days to gradually study, meditate, and apply
    the truths and commands found in this section.
Be prepared to share, encourage, and strengthen others 
    as we discuss these questions in our Friday night Bible study.

\BE
\section*{1 Peter 1:14-16}
\I  What does Peter command us \emph{not to} do,
        and what does Peter command us \emph{to} do in these verses?
\I  What is the fundamental reason why we are to obey this command?
    How does this relate to being his obedient children (v. 14)?
    See also Leviticus 11:44-45, 19:2, 20:7.
\I  Application:
    \BE
    \I  Identify the area of your life which continues most to be conformed
            to a past pattern of sin.
    \I  What is the root desire (i.e. lust) that drives you to sin in this way?
    \I  Consider Jesus, who was tempted in the same way (Hebrews 4:15).
        Search the Scripture and find how He overcame such temptation, 
            give verses.
    \I  What is the one change in your life 
            with which you can most effectively overcome this temptation?
        Consider Jesus' example from your previous answer.
    \I  Write down one Scripture that you will memorize and use to 
            fight temptation when tempted in this area.
    \EE
\section*{1 Peter 1:17-21}
\I  What is the command that we are given in these verses?
    Explain what it means, and write down the command using your own words.
\I  What are the two motivations Peter gives us, to help us to obey the command?
\I  Why is Jesus' blood so precious? Give verses.
\I  Read 1 Corinthians 3:11-15, which speaks of God's judging of believers.
    \BE
    \I  Prayerfully evaluate your Christian life up to this point.
        In what areas have you built with gold, silver, and precious stones?
        In what areas have you built with wood, hay, and straw?
    \I  Choose an area that you feel would be burned up in the fires of testing.
        Write down a verse to remember that teaches what God's will for you is
            in this area of your life.
    \EE

\section*{1 Peter 1:22-25}
\section*{1 Peter 2:1-10}
\section*{1 Peter 2:11-12}
\EE
\end{document}