### 2012-01-21 sermons/Revelation1:1-3.tex
Started Rev1:1-3 sermon.

\documentclass{article}
\newcommand{\BI}{\begin{itemize}}
\newcommand{\EI}{\end{itemize}}
\newcommand{\I}{\item}
\begin{document}
\part*{Revelation 1:1-3 --- The Revelation of Jesus Christ}
\BI
\section*{Notes}
\I  The point of this sermon is:
        \begin{quote} 
        You need to know and obey the Revelation of Jesus Christ.
        \end{quote}

This Scripture prepares us to read and obey the rest of the book.
It does so by telling us what Revelation is about
    and what we must do with it.

In this light, the goal of this sermon is to place a burden on the hearers 
    to know and become familiar with the contents of this book, 
    and to heed the commands given in it.

\section*{Introduction}
\section{The Nature of the Revelation of Jesus Christ}
\subsection{What is it about?}
\I The subject of Revelation is Jesus Christ.
\subsection{Who is it from?}
\subsection{What is its purpose?}
\subsection{Who is it for?}
\subsection{What is its focus?}

\section{The Authority of the Revelation of Jesus Christ}
\subsection{The word of God}
\subsection{The testimony of Jesus Christ}
\subsection{All things that he saw}

\section{The Blessing of the Revelation of Jesus Christ}
\EI
\end{document}