### 2012-01-27 sermons/Revelation1:1-3.tex
Updated Rev 1:1-3 sermon.

\documentclass{article}
\newcommand{\BI}{\begin{itemize}}
\newcommand{\EI}{\end{itemize}}
\newcommand{\I}{\item}
\newcommand{\Q}[1]{\begin{quote} #1 \end{quote}}
\begin{document}
\part*{Revelation 1:1-3 --- The Revelation of Jesus Christ}
\tableofcontents
\BI
\section*{Notes}
\I  The point of this sermon is:
        \begin{quote} 
        You need to know and obey the Revelation of Jesus Christ.
        \end{quote}

This Scripture prepares us to read and obey the rest of the book.
It does so by telling us what Revelation is about
    and what we must do with it.

In this light, the goal of this sermon is to place a burden on the hearers 
    to know and become familiar with the contents of this book, 
    and to heed the commands given in it.

\section*{Introduction}
\section{The Nature of the Revelation of Jesus Christ}
\subsection{Who does it reveal?}
\I The book of Revelation reveals Jesus Christ.
\subsubsection{He is God}
\I Rev 1:11 \Q{``I am the Alpha and the Omega, the First and the Last.''}
\I Rev 1:17 \Q{``Do not be afraid, I am the First and the Last.''}
\I Rev 22:13 \Q{``I am the Alpha and the Omega, the Beginning and the End, 
                the First and the Last.''}
\I Rev 1:13 \Q{\ldots and in the midst of the seven lampstands One 
                like the Son of Man, \ldots}
\I Rev 3:1, 5:6
            \Q{``These things says He who has the seven Spirits of God 
                and the seven stars: \ldots''}    
            \Q{And I looked, and behold, in the midst of the throne 
                and of the four living creatures, 
                and in the midst of the elders, 
                stood a Lamb as though it had been slain, 
                having seven horns and seven eyes, 
                which are the seven Spirits of God sent out into all the earth.}
\I Rev 3:14 \Q{``These things says the Amen, the Faithful and True Witness,
                the Beginning of the creation of God; \ldots}
\I Rev 22:16 \Q{``I, Jesus, have sent My angel to testify to you these things 
                in the churches, I am the Root and the Offspring of David, 
                the Bright and Morning Star.''}
\I Rev 1:16 \Q{\dots and His countenance was like the sun shining 
                in its strength.}

\subsubsection{He is the Savior}
\I Rev 1:5 \Q{To Him who loved us and washed us from our sins 
                in His own blood, \ldots}
\I Rev 1:18 \Q{I am He who lives, and was dead, and behold, 
                I am alive forevermore.}
\I Rev 5:6 \Q{And I looked, and behold, in the midst of the throne
                and of the four living creatures, and in the midst of the
                elders, stood a Lamb as though it had been slain, \ldots}
\I Rev 13:8 \Q{All who dwell on the earth will worship him [the beast], 
                whose names have not been written in the Book of Life
                of the Lamb who was slain from the foundation of the world.}
\I Rev 5:9 \Q{``You are worthy to take the scroll, \\
                And to open its seals; \\
                For You were slain, \\
                And have redeemed us to God by Your blood \\
                Out of every tribe and tongue and people and nation, \ldots''}
\I Rev 5:12 \Q{``Worthy is the Lamb who was slain \\
                To receive power and riches and wisdom, \\
                And strength and honor and glory and blessing!''}
\I Rev 5:13 \Q{``Blessing and honor and glory and power \\
                Be to Him who sits on the throne, \\
                And to the Lamb, forever and ever!''}
\I Rev 5:5 \Q{``Do not weep. Behold, the Lion of the tribe of Judah,
                the Root of David, has prevailed to open the scroll 
                and to loose its seven seals.''}
\I Rev 3:7 \Q{``He who has the key of David, He who opens and no one shuts, 
                and shuts and no one opens.''}
\I Rev 1:18 \Q{``And I have the keys of Hades and of Death.''}
\I Rev 19:7 \Q{``Let us be glad and rejoive and give Him glory,
                for the marriage of the Lamb has come,
                and His wife has made herself ready.''}
\I Rev 7:17 \Q{``for the Lamb who is in the midst of the throne 
                will shepherd them and lead them to living fountains of waters.
                And God will wipe away every tear from their eyes.''}
\I Rev 21:22-23 \Q{But I saw no temple in it, for the Lord God Almighty 
                and the Lamb are its temple. The city had no need 
                of the sun or of the moon to shine in it, 
                for the glory of God illuminated it. The Lamb is its light.}
\I Rev 22:1-2 \Q{And he showed me a pure river of water of life, 
                clear as crystal, proceeding from the throne of God 
                and of the Lamb. In the middle of its street, and on either
                of the river, was the tree of life, which bore twelve fruits,
                each tree yielding its fruit every month. 
                The leaves of the tree were for the healing of the nations.}
\I Rev 22:3-4 \Q{And there shall be no more curse, but the throne of God
                and of the Lamb shall be in it, and His servants shall 
                serve Him. They shall see His face, and His name shall be on
                their foreheads.}

\subsubsection{He is King}
\subsection{Who is it from?}
\subsection{What is its purpose?}
\subsection{Who is it to?}
\subsection{What is its focus?}

\section{The Authority of the Revelation of Jesus Christ}
\subsection{The word of God}
\subsection{The testimony of Jesus Christ}
\subsection{All things that he saw}

\section{The Blessing of the Revelation of Jesus Christ}
\EI
\end{document}