### 2012-02-02 sermons/Revelation1:1-3.tex
Updated sermon.

\documentclass{article}
\newcommand{\BI}{\begin{itemize}}
\newcommand{\EI}{\end{itemize}}
\newcommand{\I}{\item}
\newcommand{\Q}[1]{\begin{quote} #1 \end{quote}}
\begin{document}
\part*{Revelation 1:1-3 --- The Revelation of Jesus Christ}
\tableofcontents
\BI
\section*{Notes}
\I  The point of this sermon is:
        \begin{quote} 
        You need to know and obey the Revelation of Jesus Christ.
        \end{quote}

This Scripture prepares us to read and obey the rest of the book.
It does so by telling us what Revelation is about
    and what we must do with it.

In this light, the goal of this sermon is to place a burden on the hearers 
    to know and become familiar with the contents of this book, 
    and to heed the commands given in it.

\section*{Introduction}
\section{The Nature of the Revelation of Jesus Christ}
\subsection{Who does it reveal?}
\I The book of Revelation reveals Jesus Christ.
\subsubsection{He is God}
\I Rev 1:11 \Q{``I am the Alpha and the Omega, the First and the Last.''}
\I Rev 1:17 \Q{``Do not be afraid, I am the First and the Last.''}
\I Rev 22:13 \Q{``I am the Alpha and the Omega, the Beginning and the End, 
                the First and the Last.''}
\I Rev 1:13 \Q{\ldots and in the midst of the seven lampstands One 
                like the Son of Man, \ldots}
\I Rev 3:1, 5:6
            \Q{``These things says He who has the seven Spirits of God 
                and the seven stars: \ldots''}    
            \Q{And I looked, and behold, in the midst of the throne 
                and of the four living creatures, 
                and in the midst of the elders, 
                stood a Lamb as though it had been slain, 
                having seven horns and seven eyes, 
                which are the seven Spirits of God sent out into all the earth.}
\I Rev 3:14 \Q{``These things says the Amen, the Faithful and True Witness,
                the Beginning of the creation of God; \ldots}
\I Rev 22:16 \Q{``I, Jesus, have sent My angel to testify to you these things 
                in the churches, I am the Root and the Offspring of David, 
                the Bright and Morning Star.''}
\I Rev 1:16 \Q{\dots and His countenance was like the sun shining 
                in its strength.}

\subsubsection{He is the Savior}
\I Rev 1:5 \Q{To Him who loved us and washed us from our sins 
                in His own blood, \ldots}
\I Rev 1:18 \Q{I am He who lives, and was dead, and behold, 
                I am alive forevermore.}
\I Rev 5:6 \Q{And I looked, and behold, in the midst of the throne
                and of the four living creatures, and in the midst of the
                elders, stood a Lamb as though it had been slain, \ldots}
\I Rev 13:8 \Q{All who dwell on the earth will worship him [the beast], 
                whose names have not been written in the Book of Life
                of the Lamb who was slain from the foundation of the world.}
\I Rev 5:9 \Q{``You are worthy to take the scroll, \\
                And to open its seals; \\
                For You were slain, \\
                And have redeemed us to God by Your blood \\
                Out of every tribe and tongue and people and nation, \ldots''}
\I Rev 5:12 \Q{``Worthy is the Lamb who was slain \\
                To receive power and riches and wisdom, \\
                And strength and honor and glory and blessing!''}
\I Rev 5:13 \Q{``Blessing and honor and glory and power \\
                Be to Him who sits on the throne, \\
                And to the Lamb, forever and ever!''}
\I Rev 5:5 \Q{``Do not weep. Behold, the Lion of the tribe of Judah,
                the Root of David, has prevailed to open the scroll 
                and to loose its seven seals.''}
\I Rev 3:7 \Q{``He who has the key of David, He who opens and no one shuts, 
                and shuts and no one opens.''}
\I Rev 1:18 \Q{``And I have the keys of Hades and of Death.''}
\I Rev 19:7 \Q{``Let us be glad and rejoive and give Him glory,
                for the marriage of the Lamb has come,
                and His wife has made herself ready.''}
\I Rev 7:17 \Q{``for the Lamb who is in the midst of the throne 
                will shepherd them and lead them to living fountains of waters.
                And God will wipe away every tear from their eyes.''}
\I Rev 21:22-23 \Q{But I saw no temple in it, for the Lord God Almighty 
                and the Lamb are its temple. The city had no need 
                of the sun or of the moon to shine in it, 
                for the glory of God illuminated it. The Lamb is its light.}
\I Rev 22:1-2 \Q{And he showed me a pure river of water of life, 
                clear as crystal, proceeding from the throne of God 
                and of the Lamb. In the middle of its street, and on either
                of the river, was the tree of life, which bore twelve fruits,
                each tree yielding its fruit every month. 
                The leaves of the tree were for the healing of the nations.}
\I Rev 22:3-4 \Q{And there shall be no more curse, but the throne of God
                and of the Lamb shall be in it, and His servants shall 
                serve Him. They shall see His face, and His name shall be on
                their foreheads.}

\subsubsection{He is the conquering King}
\I Rev 1:7 \Q{Behold, He is coming with clouds, and every eye will see Him, 
                even they who pierced Him. And all the tribes of the earth
                will mourn because of Him. Even so, Amen.}
\I Rev 1:14-16 \Q{His head and hair were white like wool, as white as snow,
                and His eyes like a flame of fire; His feet were like brass, 
                as if refined in a furnace, and His voice as the sound of 
                many waters; He had in his right hand seven stars, 
                out of His mouth went a sharp two-edged sword, 
                and His countenance was like the sun shining in its strength.}
\I Rev 6:16-17 \Q{\ldots and said to the mountains and rocks, 
                ``Fall on us and hide us from the face of Him who sits on the
                throne and from the wrath of the Lamb! For the great day of His
                wrath has come, and who is able to stand?''}
\I Rev 14:14 \Q{Then I looked, and behold, a white cloud, and on the cloud
                sat One like the Son of Man, having on His head a golden crown, 
                and in His hand a sharp sickle.}
\I Rev 17:14 \Q{``These [ten kings and the beast] will make war with the Lamb, 
                and the Lamb will overcome them, for He is Lord of lords and
                King of kings; and those who are with Him are called, 
                chosen, and faithful.''}
\I Rev 19:11-16 \Q{Now I saw heaven opened, and behold, a white horse.
                And He who sat on Him was called Faithful and True,
                and in righteousness He judges and makes war.
                His eyes were like a flame of fire, and on His head were
                many crowns. He had a name written that no one knew except
                Himself. He was clothed with a robe dipped with blood,
                and His name is called The Word of God.
                And the armies of heaven, clothed in fine linen, white and
                clean, followed Him on white horses.
                Now out of His mouth goes a sharp sword, that with it He 
                should strike the nations. And He Himself will rule them 
                with a rod of iron. He Himself treads the winepress of the
                fierceness and wrath of Almighty God. And He has on His robe 
                and on His thigh a name written:
                \Q{KING OF KINGS \\ AND LORD OF LORDS.}}

\I Rev 20:15 \Q{And anyone not found written in the Book of Life was cast 
                into the lake of fire.}
            Note: we know this is the Lamb's Book of Life from Rev 13:8.

\subsection{Who is it from?}
\subsection{What is its purpose?}
\subsection{Who is it to?}
\subsection{What is its focus?}

\section{The Authority of the Revelation of Jesus Christ}
\subsection{The word of God}
\subsection{The testimony of Jesus Christ}
\subsection{All things that he saw}

\section{The Blessing of the Revelation of Jesus Christ}
\EI
\end{document}