### 2012-02-03 scribblings/symmA.tex
Draft of symmA.

\documentclass{article}
\usepackage{amsmath}
\newcommand{\BE}{\begin{equation}}
\newcommand{\EE}{\end{equation}}
\newcommand{\D}{\Delta}
\begin{document}
\BE (AD)^T = DA^T \EE
\BE (AD)_{ij} = A_{ij} d_j \EE
\BE (DA^T)_{ij} = d_i (A^T)_{ij} = A_{ji} d_i \EE
Find $d$ such that $AD$ is symmetric where 
\BE A = \nabla \times \nabla \times \EE
for a non-regular cartesian grid.
In other words, find $d$ such that
\BE A_{ij} d_j = A_{ji} d_i, \quad \text{for all $i, j$.} \EE
\begin{quote} FIGURE HERE \end{quote}
For the figure above,
\BE A_{ba} = -\frac{1}{\D x_b \D y_b} \EE
\BE A_{ab} = -\frac{1}{\D x_a \D y_a} \EE
\BE d_i \propto \frac{1}{\D x_i \D y_i} \EE
Also, we will have
\BE A_{ca} = -\frac{1}{\D x_b \D z_b} \EE
\BE A_{ac} = -\frac{1}{\D x_a \D z_a} \EE
\BE d_i \propto \frac{1}{\D x_i \D z_i} \EE
Lastly, we already have $A_{aa'} = A_{a'a}$ 
    for $a$ and $a'$ in the same direction.

The conclusion, then, is
\BE d_i = \frac{1}{\D x_i \D y_i \D z_i} \EE

\end{document}