### 2012-02-28 presentations/lightlabsFDS_mar-2012/pres.tex
Started group presentation.

\documentclass[landscape]{foils}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[pdf]{pstricks}
\input defs.tex
\raggedright
\special{! TeXDict begin /landplus90{true}store end }
\renewcommand{\oursection}[1]{
\foilhead[-1.0cm]{#1}
}

\title{An introduction to lightlabsFDS}
\author{}
\MyLogo{}
\date{}

\begin{document}
\setlength{\parskip}{0cm}
\maketitle

\BIT \itemsep -1pt
\I  The lightlabsFDS mission
\I  Solver
\I  Interface
\I  Hardware
\I  End it
\EIT

\vfill

\oursection{The lightlabsFDS mission}
\textbf{To enable engineers and scientists to characterize optical structures 
        quickly, simply, and cost-effectively.}
\vspace{10 mm}

Innovate on three fronts to make this a reality:
\BIT
\I  Solve for electromagnetic fields in the frequency domain,
\I  Run simulations from pre-installed scientific software (Matlab), 
\I  Offload computation to centralized custom-tuned hardware.
\EIT

\oursection{Solving electromagnetics in the frequency domain}
\BIT
\I  lightlabsFDS: Frequency-Domain Solver
\I  Solves 
    \BEQ \nabla \times \mu^{-1} \nabla \times E - \omega^2 \epsilon E 
            = -i \omega J. \EEQ
    \BIT
    \I  Inputs: frequency $(\omega)$, structure $(\mu, \epsilon)$, 
            and excitation $(J)$.
    \I  Outputs: electromagnetic fields $(E, H, D, B)$.
    \EIT
\I  \emph{Many} practical advantages over existing time-domain solvers.
\EIT
\newpage

\oursection{Example: time-domain solver}

\psset{unit=1.5cm}
\psset{gridcolor=green, subgridcolor=yellow}
\begin{center}
\begin{pspicture}(8,5)
    \let\psgrid\relax

    \psframe*[linecolor=gray, fillcolor=gray, fillstyle=solid] (1,2.2)(4,2.8)
    \psframe*[linecolor=gray, fillcolor=gray, fillstyle=solid] (3,1)(5,4)
    \psframe*[linecolor=gray, fillcolor=gray, fillstyle=solid] (4,1.4)(7,2)
    \psframe*[linecolor=gray, fillcolor=gray, fillstyle=solid] (4,3.0)(7,3.6)
    \rput[B](4,4.1){Device}
    \psline[linestyle=dashed](1.6,0.6)(1.6,4.0)
    \rput[B](1.6,4.1){in}
    \psline[linestyle=dashed](6.4,0.6)(6.4,4.0)
    \rput[B](6.4,4.1){out}
\end{pspicture}
\end{center}
Time-domain issues include
\BIT
\I  Input: clean excitation at input requires an auxiliary simulation
\I  Device: approximations required for material dispersion
\I  Output: overlap integrals must be repeatedly calculated 
        \emph{during} the simulation
\EIT
\newpage

\begin{center}
\begin{pspicture}(8,5)
    \psgrid

    \psframe*[linecolor=gray, fillcolor=gray, fillstyle=solid] (1,2.2)(4,2.8)
    \psframe*[linecolor=gray, fillcolor=gray, fillstyle=solid] (3,1)(5,4)
    \psframe*[linecolor=gray, fillcolor=gray, fillstyle=solid] (4,1.4)(7,2)
    \psframe*[linecolor=gray, fillcolor=gray, fillstyle=solid] (4,3.0)(7,3.6)
    \rput[B](4,4.1){Device}
    \psline[linestyle=dashed](1.6,0.6)(1.6,4.0)
    \rput[B](1.6,4.1){in}
    \psline[linestyle=dashed](6.4,0.6)(6.4,4.0)
    \rput[B](6.4,4.1){out}
\end{pspicture}
\end{center}
\end{document}