### 2012-02-29 presentations/lightlabsFDS_mar-2012/pres.tex
Updated presentation a little.

\documentclass[landscape]{foils}
\usepackage{graphicx}
\usepackage{amsmath}
% \usepackage[pdf]{pstricks}
\input defs.tex
\raggedright
\special{! TeXDict begin /landplus90{true}store end }
\renewcommand{\oursection}[1]{
\foilhead[-1.0cm]{#1}
}

\title{An introduction to lightlabsFDS}
\author{}
\MyLogo{}
\date{}

\begin{document}
\setlength{\parskip}{0cm}
\maketitle

\BIT \itemsep -1pt
\I  The lightlabsFDS mission
\I  Solver
\I  Interface
\I  Hardware
\I  End it
\EIT

\vfill

\oursection{The lightlabsFDS mission}
\textbf{To enable engineers and scientists to characterize optical structures 
        quickly, simply, and cost-effectively.}
\vspace{10 mm}

Innovate on three fronts to make this a reality:
\BIT
\I  Solve for electromagnetic fields in the frequency domain,
\I  Run simulations from pre-installed scientific software (Matlab), 
\I  Offload computation to centralized custom-tuned hardware.
\EIT

\oursection{Solving electromagnetics in the frequency domain}
\BIT
\I  lightlabsFDS: Frequency-Domain Solver
\I  Solves 
    \BEQ \nabla \times \mu^{-1} \nabla \times E - \omega^2 \epsilon E 
            = -i \omega J. \EEQ
    \BIT
    \I  Inputs: frequency $(\omega)$, structure $(\mu, \epsilon)$, 
            and excitation $(J)$.
    \I  Outputs: electromagnetic fields $(E, H, D, B)$.
    \EIT
\I  \emph{Many} practical advantages over existing time-domain solvers.
\EIT
\newpage

\oursection{Example}

% \psset{unit=1.5cm}
% \psset{gridcolor=green, subgridcolor=yellow}
% \begin{center}
% \begin{pspicture}(8,5)
%     \let\psgrid\relax
% 
%     \psframe*[linecolor=gray, fillcolor=gray, fillstyle=solid] (1,2.2)(4,2.8)
%     \psframe*[linecolor=gray, fillcolor=gray, fillstyle=solid] (3,1)(5,4)
%     \psframe*[linecolor=gray, fillcolor=gray, fillstyle=solid] (4,1.4)(7,2)
%     \psframe*[linecolor=gray, fillcolor=gray, fillstyle=solid] (4,3.0)(7,3.6)
%     \rput[B](4,4.1){Device}
%     \psline[linestyle=dashed](1.6,0.6)(1.6,4.0)
%     \rput[B](1.6,4.1){in}
%     \psline[linestyle=dashed](6.4,0.6)(6.4,4.0)
%     \rput[B](6.4,4.1){out}
% \end{pspicture}
% \end{center}
Time-domain issues include
\BIT
\I  Input: clean excitation at input requires an auxiliary simulation
\I  Device: approximations required for material dispersion
\I  Output: overlap integrals must be repeatedly calculated 
        \emph{during} the simulation
\EIT
\newpage

% \begin{center}
% \begin{pspicture}(8,5)
%     \psgrid
% 
%     \psframe*[linecolor=gray, fillcolor=gray, fillstyle=solid] (1,2.2)(4,2.8)
%     \psframe*[linecolor=gray, fillcolor=gray, fillstyle=solid] (3,1)(5,4)
%     \psframe*[linecolor=gray, fillcolor=gray, fillstyle=solid] (4,1.4)(7,2)
%     \psframe*[linecolor=gray, fillcolor=gray, fillstyle=solid] (4,3.0)(7,3.6)
%     \rput[B](4,4.1){Device}
%     \psline[linestyle=dashed](1.6,0.6)(1.6,4.0)
%     \rput[B](1.6,4.1){in}
%     \psline[linestyle=dashed](6.4,0.6)(6.4,4.0)
%     \rput[B](6.4,4.1){out}
% \end{pspicture}
% \end{center}
\BIT
\I  Fundamental problem: trying to use a time-domain solver as a 
        frequency-domain solver.
\I  Additionally, no method to measure simulation error!
\EIT
\newpage

\BIT
\I  Allow direct access to frequency-domain data.
\I  Take care of input and output fields outside of the simulation.
\I  Material dispersion explicitly defined.
\I  Simulation error well defined.
\EIT
\newpage

\BIT
\I  Frequency-domain solver made possible by correct choice of PML and 
        linear algebra algorithm.
\I  See: Wonseok Shin, Shanhui Fan, 
            ``Choice of the perfectly matched layer boundary condition for 
            frequency-domain Maxwell's equations solvers'',
            Journal of Computational Physics (January 2012).
\EIT

\oursection{Interface}
\BIT
\I  No installation required.
\begin{verbatim}
>>> [E, H] = fds(omega, epsilon, J); % Run FDS.
\end{verbatim}
\EIT
\end{document}