### 2012-03-08 sandbox/pb_template/pbt.tex
Added prayer bulletin template (latex).

\documentclass[12pt,twocolumn]{article}
\usepackage[landscape]{geometry}
\pagestyle{empty}
\newcommand{\B}{\begin{description}}
\newcommand{\I}[1]{\item[#1]}
\newcommand{\E}{\end{description}}
\begin{document}
\subsection*{Notes}
\pagebreak
\section*{Providence Bible Church Prayer Bulletin}
\begin{quote}
    Our Father in Heaven, \\
    Hallowed be Your name. \\
    Your kingdom come. \\
    Your will be done \\
    On earth as it is in heaven.
\end{quote}
\B
\I{Easter} Pray for our ministry oportunity with Lakewood elementary.
\E
\begin{quote}
    Give us this day our daily bread, \\
    And forgive us our debts, \\
    As we forgive our debtors. \\
    And do not lead us into temptation, \\
    But deliver us from the evil one.
\end{quote}
\B
\I{Needs} the other section...
\E

\begin{quote}
    For Yours is the kingdom and the power and the glory forever. Amen.
\end{quote}
\end{document}