### 2012-03-13 bible/1Peter2:18-25.tex
Added Peter Bible study.

\documentclass{article}
\usepackage{fullpage}
\newcommand{\answerbox}{\vspace{1.5cm}}
\newcommand{\BE}{\begin{enumerate}}
\newcommand{\EE}{\end{enumerate}}
\newcommand{\I}{\item}
\begin{document}
\part*{1 Peter 2:18-25}

\BE
\I  What command are servants given and why are they to obey it? \answerbox
\I  Describe the circumstances of the submissive servant that Peter is exhorting
    his readers to be. What has be done? Why did he do it? And why is his master
    punishing him? \answerbox
\I  Name people from the Bible who patiently endured unjust suffering. 
    In what way were these commendable to God? \answerbox
\I  Who is ultimately responsible for the unjust suffering of Christians?
    \answerbox
\I  How can it be right for a holy and just God to give undeserved punishment?
    \answerbox
\I  From Jesus' example, what are you to do when suffering unjustly? \answerbox
\I  How do verses 24 and 25 describe who we were before Christ and then what
    Christ has done for us? \answerbox
\I  What perspective does this give us in undeserved suffering? \answerbox
\EE
\end{document}