### 2012-04-12 bible/1Peter3:1-7.tex
Added Peter Bible study.

\documentclass{article}
\usepackage{fullpage}
\newcommand{\answerbox}{\vspace{1.5cm}}
\newcommand{\BE}{\begin{enumerate}}
\newcommand{\EE}{\end{enumerate}}
\newcommand{\I}{\item}
\begin{document}
\part*{1 Peter 3:1-7}

\BE
\section*{Wives, v.1-6}
\I  According to the passage, why are wives to be submissive to their own
    husbands? Compare with Ephesians 5:22-24 and Colossians 3:18. \answerbox
\I  Define the following terms from verses 2 and 4: 
    chaste, fear, gentle, quiet. \answerbox
\I  What were some of the specific ways in which Sarah obeyed Abraham?
    See Genesis 12-23. \answerbox
\section*{Husbands, v.7}
\I  What principles undergird \emph{both} v.1-6 and v.7, as evident by the word 
    ``likewise'' (v.7)? \answerbox
\I  In what ways is a husband to dwell with understanding, and give honor
    to his wife? \answerbox
\I  What does the warning tell us about the importance of the command?
    In what other circumstances are a believer's prayers hindered? \answerbox
\EE
\end{document}