### 2012-04-23 sermons/Revelation1:1b-3.tex
Started on new sermon.

\documentclass[12pt]{article}
\newcommand{\BI}{\begin{itemize}}
\newcommand{\EI}{\end{itemize}}
\newcommand{\I}{\item}
\newcommand{\Q}[1]{\begin{quote} #1 \end{quote}}
\begin{document}
\part*{Revelation 1:1b-3 --- The Power and Promise of the Revelation of Jesus Christ}
\tableofcontents
\BI
\section*{Notes}
\I  The point of this sermon is:
        \begin{quote} 
        You need to know and obey the Revelation of Jesus Christ.
        \end{quote}
\I  Remember that with every subsection, your job is to answer three questions:
    \begin{enumerate}
    \I  What does it mean?
    \I  Why does it matter?
    \I  How does it change my life?
    \end{enumerate}

\section*{Introduction}
\I  John's prologue underscores the gravity and urgency of Revelation.
\I  The Person of Revelation.
\I  The Purpose of Revelation.

\section{The Power of Revelation}
\subsection{The word of God}
\subsection{The testimony of Jesus Christ}
\subsection{A personal eyewitness account}

\section{The Promise of Revelation}
\subsection{What is the blessing?}
\subsection{Who is the blessing for?}
\subsection{What does the blessing mean?}

\EI
\end{document}