### 2012-05-29 presentations/meta_book/1d.tex
Added progress on meta book chapter.

\section{1D resonator design}
We now build up to the objective-first formulation by example. 
And from a different perspective.

We're going to take a step back and start with the most naive inverse design strategy possible.
Then we'll end with objective first.
This motivates the need for the alternating directions and objective-first strategy.

\subsection{Direct solve of structure}
This is the simplest thing you can do. Basically, set $x$ and solve for $p$.
Also, take out the relaxed constraint on $p$.
\BA \minimize{p} \| B(x) p - d(x) \|^2 \EA
\myfig{1d/leastsquares}{test}
We perfectly satisfy the field but...

\subsection{Regularized solve of structure}
We add a term to try to control $p$
\BA \minimize{p} \| B(x) p - d(x) \|^2 + \eta \|p - p_0\|^2 \EA
\myfig{1d/regularized}{test}
Better, but a trade-off between field accuracy and structure variation.

\subsection{Alternating directions solve}
Now we actually do both fields.
\BA \minimize{p} \| B(x) p - d(x) \|^2 + \eta_1 \|p - p_\text{prev}\|^2 \EA

\BA \minimize{x} \| A(p) x - b(p) \|^2 + \eta_2 \|x - x_\text{prev}\|^2 \EA
\myfig{1d/complementary}{test}

\subsection{Alternating directions solve with bounded $p$}
\BA \minimize{p} \| B(x) p - d(x) \|^2 \\
    \subto 0 \le p \le 1 \EA

\BA \minimize{x} \| A(p) x - b(p) \|^2 + \eta_2 \|x - x_\text{prev}\|^2 \EA
\myfig{1d/bounded}{test}