### 2012-05-29 presentations/meta_book/my_chapter.tex

% Some custom commands.
\newcommand{\BE}{\begin{equation}}
\newcommand{\EE}{\end{equation}}
\newcommand{\BA}{\begin{eqnarray}}
\newcommand{\EA}{\end{eqnarray}}
\newcommand{\curl}{\nabla\times}
\newcommand{\minimize}[1]{\JLUminimize_{#1}\;&}
\newcommand{\subto}{\text{subject to}\;&}
\newcommand{\myfig}[2]{\begin{figure}[!h]\includegraphics[width=\textwidth]{fig/#1}\caption{#2}\label{#1}\end{figure}}

\chapter{Objective-First Nanophotonic Design}
\label{intro}

\abstract{The abstract for the book.}

Introductory paragraph.

\input{wave_eq}
\input{ob1}
\input{1d}
\input{res}
\input{wg}

\section{Metamaterials design}
\subsection{Modification of the design objective}
\subsection{Cloak devices}
\subsection{Mimic devices}

\section{Extending the method}
\subsection{3D}
\subsection{Multi-mode}
\subsection{Robustness}

\section{Appendix}
\subsection{Full 3D curl}
\subsection{1D}
\subsection{2D}
\subsection{2.5D}