### 2012-05-29 presentations/meta_book/ob1.tex

\section{The objective-first design problem}
intro

\subsection{Design objectives}
Talk about $f(x)$ and that we are interested in convex ones.

\subsection{Convexity}
Convex optimization quick intro.

\subsection{Typical design formulation}
Typically,
\BA \minimize{x,p} f(x) \\
    \subto g(x, p) = 0 \\
        & p \in {0,1} \EA

\subsection{Objective-first design formulation}
Objective-first does 
\BA \minimize{x,p} \|g(x,p)\|^2 \\
    \subto f(x) = f_\text{ideal} \\
        & 0 \le p \le 1 \EA

This is a bi-convex problem, which we solve using an alternating directions method.

\subsection{Field sub-problem}
\BA \minimize{x} \| A(p) x - b(p) \|^2 \\
    \subto f(x) = f_\text{ideal} \EA

\subsection{Structure sub-problem}
\BA \minimize{p} \| B(x) p - d(x) \|^2 \\
    \subto 0 \le p \le 1 \EA