### 2012-05-29 presentations/meta_book/wg.tex

\section{Waveguide coupler design}
Finally, we now implement 
    the full objective-first design strategy
    by doing an alternating directions solve.

\subsection{Choice of design objective}
We choose, for generality,

\BE f(x) = \begin{cases}
        x & \text{at boundary}, \\
        0 & \text{elsewhere}.
        \end{cases} \EE

    where the boundary denotes the two outermost layers of the design space.
This means

\BE f_\text{ideal} = \begin{cases}
        x_\text{perfect} & \text{at boundary}, \\
        0 & \text{elsewhere}.
        \end{cases} \EE

We then do

\BA \minimize{x} \| A(p) x - b(p) \|^2 \\
    \subto f(x) = f_\text{ideal} \EA

\BA \minimize{p} \| B(x) p - d(x) \|^2 \\
    \subto 0 \le p \le 1 \EA

\subsection{Coupling between dielectric waveguide modes}
\myfig{wg/1}{test}
\myfig{wg/2}{test}
\myfig{wg/3}{test}

\subsection{Coupling to plasmonic waveguide modes}
\myfig{wg/4}{test}
\myfig{wg/5}{test}