### 2012-05-31 presentations/meta_book/my_chapter.tex
Getting started on text.

% Some custom commands.
\newcommand{\BE}{\begin{equation}}
\newcommand{\EE}{\end{equation}}
\newcommand{\BA}{\begin{eqnarray}}
\newcommand{\EA}{\end{eqnarray}}
\newcommand{\curl}{\nabla\times}
\newcommand{\minimize}[1]{\JLUminimize_{#1}\;&}
\newcommand{\subto}{\text{subject to}\;&}
\newcommand{\myfig}[2]{\begin{figure}[!h]\includegraphics[width=\textwidth]{fig/#1}\caption{#2}\label{#1}\end{figure}}
\newcommand{\BI}{\begin{itemize}\item}
\renewcommand{\I}{\item}
\newcommand{\EI}{\end{itemize}}
\newcommand{\ER}[1]{\eqref{eq:#1}}
\chapter{Objective-First Nanophotonic Design}
\label{intro}

\abstract{The abstract for the book.}

In this chapter, we introduce an ``objective-first'' strategy 
    for designing nanophotonic devices.

General description:
\BI Only uses desired fields.
\I  Can attempt to design any linear device.
\I  Allows for non-physical fields. \EI

This strategy is unique in that it
    asks the user only for what electromagnetic field
    the device should produce,
    and then attempts to generate the design
    without further user intervention.

\input{wave_eq}
\input{ob1}
% \input{1d}
% \input{res}
% \input{wg}

\section{Metamaterials design}
\subsection{Modification of the design objective}
\subsection{Cloak devices}
\subsection{Mimic devices}

\section{Extending the method}
\subsection{3D}
\subsection{Multi-mode}
\subsection{Robustness}

\section{Appendix}
\subsection{Full 3D curl}
\subsection{1D}
\subsection{2D}
\subsection{2.5D}