### 2012-06-04 presentations/meta_book/mimic.tex

\section{Optical mimic design}
\myfig{mimic/m1}{test}
\myfig{mimic/m2}{test}
\myfig{mimic/m3}{test}
\myfig{mimic/m4}{test}
\myfig{mimic/m5}{test}
\myfig{mimic/m6}{test}