### 2012-06-04 presentations/meta_book/mimic.tex
Started on mimics.

\section{Optical mimic design}
We now apply our objective-first strategy to the design of optical mimics.

We define an optical mimic to be a linear nanophotonic device which 
    mimics the output field of another device.
In this sense optical mimics are the anti-cloak;
    whereas cloaks strive to make an object's electromagnetic presence vanish,
    mimics strive to implement an object's presence without that 
    object actually being there.

As such, the design of optical mimics provides a tantalyzing approach
    to the realization of practical metamaterial devices.
That is to say, if one can reliably produce practical optical mimics,
    then producing metamaterials (which are often based on fictitious materials)
    can be accomplished by simply producing an optical mimic of that material.

In a more general sense, 
    optical mimics is really just a recasting of the thrust of 
    the objective-first design strategy in its purest form:
    the design of a nanophotonic device based purely 
    on the electromagnetic fields one wishes to have it produce.
As such, devices which perform well-known optical functions
    (e.g. focusing, lithography) can also be designed.

\subsection{Application of the objective-first strategy}
The objective-first design of optical mimics proceeds in virtually
    an identical way to the design of optical cloaks,
    the only difference being that the output modes are 
    specifically chosen to be those which produce the desired function.
For most of the examples provided, the input illumination is still an incident plane wave.

Lastly, instead of measuring efficiency, 
    we measure the relative error of the simulated field
    against that of a perfect target field
    at a relevant plane of some distance away from the device.
The location of this plane is identified as a dotted line in the subsequent figures.

\subsection{Plasmonic cylinder mimic}
Our first mimic is simply to mimic the plasmonic cylinder 
    which we cloaked in the previous section.
\myfig{mimic/m1}{Plasmonic cylinder mimic. 
                Error: 8.1\%,
                footprint: $40 \times 120$ grid points,
                wavelength: 42 grid points.}

\subsection{Diffraction-limited lens mimic}
\myfig{mimic/m3}{Full-width-half-max at focus: 1.5 $\lambda$,
                focus depth: 100 grid points.
                Error: 12.0\%,
                footprint: $40 \times 120$ grid points,
                wavelength: 25 grid points.}
\myfig{mimic/m4}{Full-width-half-max at focus: 0.5 $\lambda$,
                focus depth: 50 grid points.
                Error: 5.6\%,
                footprint: $40 \times 120$ grid points,
                wavelength: 25 grid points.}
\myfig{mimic/m5}{Full-width-half-max at focus: 0.5 $\lambda$,
                focus depth: 150 grid points.
                Error: 1.4\%,
                footprint: $40 \times 120$ grid points,
                wavelength: 25 grid points.}

\subsection{Sub-diffraction lithographic mask}
\myfig{mimic/m6}{Error: 19.8\%}

\subsection{Perfect lens mimic}
\myfig{mimic/m2}{Perfect lens mimic.
                Error: 28.6\%}