### 2012-06-14 resume/bina/res1.tex

% LaTeX file for resume 
% This file uses the resume document class (res.cls)

\documentclass{res} 
%\usepackage{helvetica} % uses helvetica postscript font (download helvetica.sty)
%\usepackage{newcent}   % uses new century schoolbook postscript font 
\setlength{\textheight}{9.5in} % increase text height to fit on 1-page

\begin{document}

\name{Jesse Lu\\[12pt]}
\address{\texttt{jesselu@stanford.edu} \\ 66 Newell Rd. Apt. O \\ East Palo Alto, CA \\ (408) 568-9356}

\begin{resume}

\section{JOB OBJECTIVE}          
    A full-time software engineering position at Bina with starting date of November 2012 
    (upon completion of doctoral degree).

\section{EDUCATION}          
    Stanford University, PhD, Electrical Engineering, November 2012 (in progress) \\
    Stanford University, Masters of Science, Electrical Engineering, May 2012 \\
    University of California Los Angeles, Bachelor of Science, Electrical Engineering, June 2006

\section{HONORS}
    Stanford Graduate Fellowship, Stanford University, 2007 \\
    Dean's Honors List, University of California Los Angeles, five quarters

\section{EXPERIENCE}
   \vspace{-0.1in}	
   \begin{tabbing}
   \hspace{2.3in}\= \hspace{2.6in}\= \kill % set up two tab positions
    {\bf Graduate Researcher} \>Jelena Vuckovic Group \>2007-Present\\
                             \>Stanford University
   \end{tabbing}\vspace{-20pt}      % suppress blank line after tabbing
    Existing nanophotonic design process consisted of trial-and-error
        where each simulation about 24 hours to complete.
    Implemented FDTD and FDFD simulation software on Nvidia GPUs to bring 
        simulation times down to 10 minutes.
    Also developed a novel ``objective-first'' nanophotonic design algorithm
        and demonstrated automated design of 
        small-footprint, high-efficiency waveguide couplers.
   \begin{tabbing}
   \hspace{2.3in}\= \hspace{2.6in}\= \kill % set up two tab positions
    {\bf Undergraduate Researcher} \>Eli Yablonovitch Group \> 2005-2006\\
                          \>University of California, Los Angeles
   \end{tabbing}\vspace{-20pt}
    Surface plasmon focusing device required a grating coupler to inject energy
        into the surface plasmon mode.
    Wrote a Matlab program interfacing with a FEM package (COMSOL),
        developed a hierarchal optimization routine as well as an extended 
        transfer matrix model to produce first-ever thoroughly optimized 
        grating coupler design.
   \begin{tabbing}%
   \hspace{2.3in}\= \hspace{2.6in}\= \kill % set up two tab positions          
   {\bf Research Assistant Intern}  \>Inko Pellicles \> Summer 2003\\
                          \>Sunnyvale, CA
   \end{tabbing}\vspace{-20pt}
    Pellicle removal left residue on expensive lithography masks.
    Assisted in researching and developing a UV irradition process
        to modify the adhesive properties of the glue.

\section{PROGRAMMING}          
    Fluent in Python, CUDA, and Matlab. \\
    Currently working with Git and Amazon EC2.

\section{PUBLICATIONS}          
    ``Objective-first design of high-efficiency, small-footprint couplers between arbitrary nanophotonic waveguide modes'' Jesse Lu, Jelena Vuckovic (Optics Express, 2012). \\
    ``Inverse design of a three-dimensional nanophotonic resonator'' Jesse Lu, Stephen Boyd, Jelena Vuckovic (Optics Express, 2011). \\
    ``Inverse design of nanophotonic structures using complementary convex optimization'' Jesse Lu, Jelena Vuckovic (Optics Express, 2010). \\
    ``Numerical optimization of a grating coupler for the efficient excitation of surface plasmons at an Ag-SiO2 interface'' Jesse Lu, Csaba Petre, Josh Conway, Eli Yablonovitch (JOSA B, 2007).

\end{resume}
\end{document}