### 2012-07-09 presentations/meta_book/my_chapter.tex

% Some custom commands.
\newcommand{\BE}{\begin{equation}}
\newcommand{\EE}{\end{equation}}
\newcommand{\BA}{\begin{eqnarray}}
\newcommand{\EA}{\end{eqnarray}}
\newcommand{\curl}{\nabla\times}
\newcommand{\minimize}[1]{\JLUminimize_{#1}\;&}
\newcommand{\subto}{\text{subject to}\;&}

% Custom figures.
\newcommand{\myfig}[2]{\begin{figure}[!h]\includegraphics[width=0.95\textwidth]{fig/#1.jpg}\caption{#2}\label{fig:#1}\end{figure}}

\newcommand{\BI}{\begin{itemize}\item}
\renewcommand{\I}{\item}
\newcommand{\EI}{\end{itemize}}

\newcommand{\ER}[1]{\eqref{eq:#1}}
\newcommand{\SR}[1]{Section~\ref{sec:#1}}
\newcommand{\sR}[1]{section~\ref{sec:#1}}
\newcommand{\FR}[1]{Figure~\ref{fig:#1}}
\newcommand{\fR}[1]{figure~\ref{fig:#1}}
\chapter{Objective-First Nanophotonic Design}
\label{intro}

\abstract{We introduce an ``objective-first'' strategy 
    for designing nanophotonic devices,
    and we demonstrate the design of nanophotonic
    coupler, cloak, and mimic devices.}

Our initial foray into design methods for nanophotonic devices
    began with a very simple and naive question: 
    could we make an inverse solver which,
    when given the electromagnetic fields we desire,
    returns the nanophotonic structure that will produce them\cite{opex3}?
In other words, since we already know how to solve for $E$ and $H$ 
    in Maxwell's equations, why can't we solve for $\epsilon$
    or even $\mu$ instead?

Not surprisingly, it did not take us long to find that such
    a simple strategy would inevitably run into many problems.

Over the subsequent years, we were able to come up with a better solution,
    which we call an ``objective-first'' strategy for nanophotonic design,
    and which we present in this chapter.
Although it is much more advanced than our original idea,
    objective-first design still carries the same fundamental concept,
    which is to specify the electromagnetic fields, and then
    to solve for a structure to produce them.

In this chapter, we present the simple theoretical underpinnings of 
    objective-first design in the first two sections,
    and then show examples of the method in action in the rest of the chapter.
We also include the source code that was used to generate
    all results presented herein\cite{code}.

\input{wave_eq}
\input{ob1}
\input{wg}
\input{cloak}
\input{mimic}
\input{extensions}

\section{Conclusion}
We have introduced an objective-first approach
    to the design of nanophotonic components, and
    applied it to the design of waveguide couplers,
    optical cloaks, and optical mimics.
In doing so, we hope to have exhibited both the simplicity
    and the breadth of our method to the design of a broad
    class of linear, single-mode devices.
In addition to posting the source code for all the examples online\cite{code}, 
    we have outlined the necessary extensions to our method in order
    to design practical, three-dimensional devices.

\section{Acknowledgements}
This work has been supported by the 
    AFOSR MURI for Complex and Robust On-chip Nanophotonics 
    (Dr. Gernot Pomrenke), grant number FA9550-09-1-0704; 
    and the Alexander von Humboldt Foundation.

\begin{thebibliography}{99.}%
\bibitem{boyd} S. Boyd, and L. Vandenberghe, 
    \emph{Convex Optimization} 
    (Cambridge University Press, 2004)
\bibitem{cvx} M. Grant and S. Boyd, \emph{CVX: Matlab software for disciplined convex programming}, version 1.21. \verb+http://cvxr.com/cvx+, January 2011.
\bibitem{admm} S. Boyd, N. Parikh, E. Chu, B. Peleato, and J. Eckstein, 
    ``Distributed Optimization and Statistical
    Learning via the Alternating Direction
    Method of Multipliers,''
    Foundations and Trends in Machine Learning, \textbf{3} 1–122 (2011).
\bibitem{pml} S. G. Johnson, ``Notes on Perfectly Matched Layers (PMLs),''
    http://math.mit.edu/~stevenj/18.369/pml.pdf, (2010). 
\bibitem{opex1} J. Lu, and J. Vuckovic, 
    ``Inverse design of nanophotonic structures using complementary convex optimization,''
    Opt. Express \textbf{18}, 3793--3804 (2011). 
% \bibitem{opex2} J. Lu, S. Boyd, and J. Vuckovic, 
%     ``Inverse design of a three-dimensional nanophotonic resonator,''
%     Opt. Express \textbf{19}, 10563--10750 (2011). 
\bibitem{opex3} J.Lu, and J. Vuckovic,
    ``Objective-first design of high-efficiency, small-footprint couplers between arbitrary nanophotonic waveguide modes,''
    Opt. Express \textbf{20}, 7221--7236 (2012).
\bibitem{code} https://github.com/JesseLu/objective-first
\bibitem{miller} O. D. Miller, ``Photonic Design: From Fundamental Solar Cell Physics to Computational Inverse Design,''
    U. C. Berkeley, (2012).
\bibitem{shin} W. Shin, S. Fan, ``Choice of the perfectly matched layer boundary condition for frequency-domain Maxwell’s equations solvers,''
    J. Comp. Phys. \textbf{231}, 3406--3431 (2012).
\bibitem{simp} M. P. Bendsoe, and O. Sigmund, 
    ``Material interpolation schemes in topology optimization,''
    Archive of Applied Mechanics \textbf{69}, 635--654 (1999).
\bibitem{taflove} A. Taflove, and S. C. Hagness, 
    \emph{Computational Electrodynamics, Third Ed.}
    Section 3.3.2, (Artech House, 2005).
\bibitem{yee} K. S. Yee,
    ``Numerical Solution of Initial Boundary Value Problems Involving
    Maxwell's Equations in Isotropic Media,''
    IEEE Trans. Antennas and Propagation \textbf{14}, 802--807 (1966).
\end{thebibliography}