### 2012-07-18 scribblings/plan.tex
Started a plan.

\documentclass{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\input defs.tex
\newcommand{\inv}[1]{\frac{1}{#1}}
\newcommand{\tC}{\tilde{C}}
\newcommand{\BSE}{\begin{subequations}\begin{align}}
\newcommand{\EA}{\end{align}}
\newcommand{\ESE}{\end{subequations}\end{align}}
\newcommand{\env}[2]{\begin{#1}#2\end{#1}}
\newcommand{\E}[1]{\begin{subequations}\begin{align}#1\end{align}\end{subequations}}
\DeclareMathOperator*{\minimize}{\text{minimize}\quad}
\newcommand{\subto}{\text{subject to}\quad}
\title{Objective-First Nanophotonic Design Plan} 
\author{Jesse Lu}
\begin{document}
\maketitle
\tableofcontents

\section{Goal}
    Software to solve the following general inverse design problem in nanophotonics.
    \E{ \minimize&  f(x) + g(z) \\
        \subto&     A(z)x - b(z) = 0}
% \end{align}\end{subequations}
    The software will include two 
\end{document}