### 2012-07-18 resume/lytro/cover_letter.tex

\documentclass[12pt]{article}
\usepackage{fullpage}
\addtolength{\textheight}{2cm}

\newcommand{\B}{\begin{description}}
\newcommand{\I}[1]{\item[#1]}
\newcommand{\E}{\end{description}}

\title{Jesse Lu}
\date{}

\begin{document}
% \maketitle
\thispagestyle{empty}
\part*{\center Jesse Lu}
\section*{Who am I? A problem solver.}
Not every problem, not all problems, and not by myself;
    but at the heart of it, solving problems is what I do well,
    it's what I enjoy doing, and it's what I would like to do at
    Lytro.

I want to join Lytro in solving technically-challenging, world-changing problems.

\section*{A solver: technologist, innovator, and communicator.}
\subsection*{\emph{Technologist}: has a deep proficiency in 
                            a unique set of technical areas.}
\B  \I{GPUs} I have developed multiple electromagnetic simulation packages
                for the GPU, the latest being a multi-GPU, dynamically-compiled
                electromagnetic solver based on CUDA.

    \I{Physics} I did my PhD in nanophotonics and have a deep expertise 
                in electromagnetics, as well as a general proficiency
                in all things scientific.

    \I{Optimization} During that PhD I developed an optimization paradigm
                for nanophotonic structures, and understood the power of
                convex optimization and related techniques.
\E

\subsection*{\emph{Innovator}: finds and combines the technical pieces needed 
                            to create something amazing.}
\B  \I{Learns fast} I intentionally expose myself to other disciplines
                in order to ``connect the dots'' in a powerful way, 
                and I quickly incorporate their salient ideas into my own projects.

    \I{Fails fast} When you innovate, you suceed only after many failures.
                I wittle down new features/methods into
                their simplest viable test cases in order 
                to fail as fast as possible.
\E

\subsection*{\emph{Communicator}: disseminates the results of innovation and 
                            engages others in furthering innovation.}
\B  \I{Public speaking} Stand me up in front of a crowd, and I'll love it.
                One of my passions is to be able to bring others to that
                ``Aha!'' moment, especially concerning complex
                technical subjects.

    \I{Connecting} Or put my in a room with a single person, 
                and I'll love that too.
                I am always eager to learn and I've found that questioning and
                exploring someone else's knowledge and perspectives is often 
                the best way to do so.
\E

In conclusion, I believe that great companies are simply a collection of great problem-solvers
    tackling great problems and I hope to join Lytro in doing 
    so\footnote{\texttt{jesselu@stanford.edu, (408)568-9356, East Palo Alto, CA.}}.
\end{document}