### 2012-07-19 sermons/Revelation1:4.tex
Started new sermon.

\documentclass[12pt]{article}
\newcommand{\BI}{\begin{itemize}}
\newcommand{\EI}{\end{itemize}}
\newcommand{\I}{\item}
\newcommand{\Q}[1]{\begin{quote} #1 \end{quote}}
\begin{document}
\part*{Revelation 1:4 --- Grace and Peace from God, Part 1}
\tableofcontents
\BI
\section*{The Need for Grace and Peace}
\I one

\section*{The Source of Grace and Peace}
\I two
\EI

\end{document}