### 2012-08-30 bible/Psalm1.tex
Added Psalm 1 notes.

\documentclass{article}
\newcommand{\BI}{\begin{itemize}\item}
\newcommand{\I}{\item}
\newcommand{\EI}{\end{itemize}}
\title{Psalm 1}

\begin{document}
\maketitle

\section{Summary of the text}
\subsection{What does it say?}
Psalm 1 contrasts the righteous and the wicked by
    \BI identifying who they are,
    \I  describing how they will live, and
    \I explaining why the contrast is so great. \EI
\subsection{Why is it there?}
This psalm is intentionally placed first
    in order to prepare us to worship God
    by first bringing us to repentance.
\subsection{How does it matter?}
Psalm 1 plays a critical role in the believers life in
    \BI discerning whether we are living 
            the life of the righteous or the wicked,
    \I  increasing our desire to live righteously, and
    \I  teaching how to do so. \EI

\end{document}