### 2012-10-19 presentations/maxwell_documentation/pres.tex
Started another group presentation.

\documentclass[landscape]{foils}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\input defs.tex
\raggedright
\special{! TeXDict begin /landplus90{true}store end }
\renewcommand{\oursection}[1]{
\foilhead[-1.0cm]{#1}
}

\title{Maxwell: bringing cloud-powered electromagnetic simulations to Matlab}
\author{}
\MyLogo{Jesse Lu, Jelena Vuckovic group, Stanford University}
\date{}

\begin{document}
\setlength{\parskip}{0cm}
\maketitle

\BIT \itemsep -1pt
\item Quick-start
\item How Maxwell works
\item How the frequency-domain solver works
\item Examples
\EIT

\vfill

\oursection{Problem Statement}
% \oursection{Plan}
% \BIT
% \item Software to solve $A^{-1}z$.
% \item Adjoint
% \EIT
\end{document}