### 2012-10-19 presentations/maxwell_documentation/pres.tex
Format of group presentation set.

\documentclass[landscape]{foils}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\input defs.tex
\raggedright
\special{! TeXDict begin /landplus90{true}store end }
\renewcommand{\oursection}[1]{
\foilhead[-1.0cm]{#1}
}

\title{Maxwell: bringing cloud-powered electromagnetic simulations to Matlab}
\author{}
\MyLogo{Jesse Lu, Jelena Vuckovic group, Stanford University}
\date{}

\begin{document}
\setlength{\parskip}{0cm}
\maketitle

\BIT \itemsep -1pt
\item Quick-start
\item How Maxwell uses the cloud (EC2)
\item How Maxwell solves electromagnetics
\item Examples
\EIT

\vfill

\oursection{Quick-start}
\oursection{How Maxwell uses the cloud (EC2)}
\oursection{How Maxwell solves electromagnetics}
\oursection{Examples}

\end{document}