### 2013-03-18 sandbox/pb_cleanup/cleanup.tex
Added cleaning schedule.

\documentclass[12pt]{article}
\usepackage{fullpage}
\title{PBC Cleaning Schedule}
\date{2013}
\newcommand{\BI}{\begin{itemize}\item}
\newcommand{\I}{\item}
\newcommand{\EI}{\end{itemize}}
\begin{document}
\maketitle
\section{What is this?}
A simple plan to be faithful to maintain the property that God has given us.
Pairs of members are assigned a specific month in which
    to clean the main church building.
Cleanings should occur \textbf{the week leading up to the 4th Sunday} 
    of the given month 
    (in order to be best spaced from the work day cleaning).
Pairs are assigned so that this time can also be used as a time of fellowship.

\section{What needs to be cleaned?}
\BI Restrooms
    \BI Toilet paper available in each stall
    \I  Paper towel dispensers filled
    \I  Soap dispensers filled
    \I  Toilets flushed and presentable \EI
\I  Pews
    \BI Bibles and hymnals straightened
    \I  Pens in each alotted space
    \I  Guest cards ($\sim2$) in each alotted space \EI
\I  Nursery and Training Room
    \BI Toys in appropriate place (e.g. bins)
    \I  Chairs straightened
    \I  Bibles and hymnals available on Nursery end table \EI
\I  Plants
    \BI Watered (only for those which are real)
    \I  Plants near pulpit look good and presentable \EI
\EI
\section{Assignments}
\BI These dates are for 2013.
\I  Date ranges denote Monday to Saturday of the week leading 
        up to the fourth Sunday of the month.
    This was chosen in order to space this cleaning between
        successive word day cleanings. \EI
\begin{center}
\begin{tabular}{|c|c|}
    \hline
    Date range & Assignment \\
    \hline \hline
    April 22-27 & Davis \& Hardin \\ \hline
    May 20-25 & Lu \& Mixsell \\ \hline
    June 17-22 & Owens \& Sheldon \\ \hline
    July  22-27 & Sisson \& Torrefranca \\ \hline
    August 19-24 & IBBC \& Davis \\ \hline
    September 23-28 & Hardin \& Lu \\ \hline
    October 21-26 & Mixsell \& Owens \\ \hline
    November 18-23 & Sheldon \& Sisson \\ \hline
    December 16-21 & Torrefrance \& IBBC \\ \hline
\end{tabular}
\end{center}
\end{document}