### 2013-05-21 presentations/fdtd_gpu/pres.tex
Added fdtd gpu presentation

\documentclass[landscape]{foils}
\usepackage{graphicx}
\usepackage{amsmath}

\raggedright
\special{! TeXDict begin /landplus90{true}store end }

\newcommand{\oursection}[1]{
\foilhead[-1.0cm]{#1}
}

\newcommand{\myfig}[1]{
    \begin{figure}[!h]
    \centerline{\includegraphics[height=6.5cm]{fig/#1.jpg}}
    \end{figure}
}
\newcommand{\widefig}[1]{
    \begin{figure}[!h]
    \centerline{\includegraphics[width=\textwidth]{fig/#1.jpg}}
    \end{figure}
}
\newcommand{\tallfig}[1]{
    \begin{figure}[!h]
    \centerline{\includegraphics[height=0.8\textheight]{fig/#1.jpg}}
    \end{figure}
}
\newcommand{\medfig}[1]{
    \begin{figure}[!h]
    \centerline{\includegraphics[height=10cm]{fig/#1.jpg}}
    \end{figure}
}
\newcommand{\bigfig}[1]{
    \begin{figure}[!h]
    \centerline{\includegraphics[height=12cm]{fig/#1.jpg}}
    \end{figure}
}
\newcommand{\BI}{\begin{itemize}\item}
\newcommand{\I}{\item}
\newcommand{\EI}{\end{itemize}}

\title{FDTD on the GPU}
\author{Jesse Lu}
\MyLogo{Jesse Lu, Jelena Vuckovic group, Stanford University}
\date{May 21, 2013}

\begin{document}
\setlength{\parskip}{0cm}
\maketitle

\begin{itemize} \itemsep -1pt
\I  The GPU today
\I  CPU vs GPU 
\I  Programming the GPU
\I  Implementing FDTD on the GPU
\I  Optimizing FDTD on the GPU
\EI

\vfill
\newpage

\oursection{The GPU today}
\BI GPU (Graphical Processing Unit) once only for 3D gaming, now everywhere.
\I  Present in virtually all computing form factors
    \BI Raspberry Pi: 24 GFLOPS Broadcom GPU
    \I  Titan Supercomputer: 20 PetaFLOPS, 18K+ Nvidia GPUs \EI
\I  Plays key role in numerous applications
    \BI graphics: games, video encoding/decoding, \ldots
    \I  science: protein folding, FDTD simulations, \ldots
    \I  misc.: bitcoin mining, password cracking, \ldots \EI
\vfill

\oursection{CPU vs. GPU}
\widefig{comp}
\newpage
% \I  Each of the GPU's 15 scalar multiprocessors contains 196 cores 
%     (almost 3000 total cores!)
% \tallfig{smx-block}
\begin{tabular}{r|c c}
     & CPU (Intel Xeon E7) & GPU (Nvidia Kepler) \\ \hline
    Number of cores & 10 & 2940 \\
    Cache size & 30MB & $\sim$2MB \\
    FLOPS (single-precision) & 0.67 TFLOPS & 3.7 TFLOPS \\
    FLOPS (double-precision) & 0.33 TFLOPS & 1.2 TFLOPS \\
    Memory bandwidth & 102 GB/s & 250 GB/s \\
    Power draw & 130 W & 250 W
\end{tabular}
\I  Spec comparison needs to always be taken with a grain of salt
    (e.g. ``theoretical maximums'')
\I  Question: What kind of workloads are most suitable for the CPU? the GPU?

\oursection{Programming the GPU}
\I  Will focus on Nvidia GPUs and the CUDA programming language
\I  CUDA uses the SIMD (Single Instruction Multiple Data) model for parallelization
\I  Programming model mirrors GPU architecture

\begin{tabular}{c c c}
Computational & Memory & Software \\ \hline\hline
core & registers & thread \\ \hline
multiprocessor & cache/shared memory & thread block \\ \hline
GPU & global memory & block grid
\end{tabular}
\oursection{Implementing FDTD on the GPU}
\I  We'll just look at a 1D $E$-field update:
    \begin{equation}
    E^i = c_0 E^i + c_1 (H^i - H^{i-1}) 
    \end{equation}
\I  This is how we would normally do it on the CPU
    \begin{verbatim}
    for (int i=0 ; i < N ; i++) {
        E[i] = c0 * E[i] + c1 * (H[i] - H[i-1]); 
    }
    \end{verbatim}

\newpage
\I  The simplest CUDA version of the update would be:
    \begin{verbatim}
    E_update<<<1, N>>>();
    ...

    __global__ update_E () {
        i = threadId;
        E[i] = c0 * E[i] + c1 * (H[i] - H[i-1]); 
    }
    \end{verbatim}
\I  Each cell is updated by a single thread
\I  All threads grouped in a single thread block

\newpage
\I  Launching multiple blocks will allow us to utilize
        multiple multiprocessors
    \begin{verbatim}
    E_update<<<N/16, 16>>>();
    ...

    __global__ update_E () {
        i = blockId * 16 + threadId;
        E[i] = c0 * E[i] + c1 * (H[i] - H[i-1]); 
    }
    \end{verbatim}

\oursection{Optimizing FDTD on the GPU}
\I  Basic optimization question to answer:
    Are we limited by the memory-bandwidth or by computational power?
\I  Optimizing memory access patterns is often the key to faster GPU codes
\I  \emph{Shared memory} resides at the multiprocessor level,
        and quickly accessed by all cores within the multiprocessor
\I  We can optimize memory access by eliminating redundant $H$-field loads

\newpage
\I  Shared memory version of the update function
    \begin{verbatim}
    E_update<<<N/16, 16>>>();
    ...

    __global__ update_E () {
        i = blockId * 16 + threadId;
        s_i = threadId;
        __shared__ float s_H[17];
        s_H[s_i+1] = H[i];
        if (threadId == 0) { s_H[0] = H[i-1] };
        __syncthreads();
        E[i] = c0 * E[i] + c1 * (s_H[s_i+1] - H[s_i]); 
    }
    \end{verbatim}

\EI
\end{document}