### 2013-08-10 sandbox/pbc-workday/2013-08.tex
August workday.

\documentclass[12pt]{article}
\newcommand{\BI}{\begin{enumerate}\item}
\newcommand{\I}{\item}
\newcommand{\EI}{\end{enumerate}}
\title{Work Day}
\begin{document}
\maketitle
Items are listed in order of \emph{approximate} priority. 
\BI Main Sanctuary
    \BI Vacuum pews and carpet
    \I  Restock visitor cards and pencils into pews.
    \I  General cleaning (remove old bulletins...) \EI
\I  Bathrooms: Regular cleaning and restocking
\I  Annex carpet estimate 
    \BI We haven't committed to carpeting the Annex, 
        but it would be really beneficial to have a cost/work/style estimate.
    \I  If unexperienced with carpeting rooms, you may want to consult someone
        who does or the internet first.
    \I  Determine amount of carpeting needed to cover annex dining room.
    \I  Visit Lowe's (811 E Arques Ave) and choose around 3 carpet samples.
    \I  Also, obtain cost estimates for each of the samples.
    \I  Hopefully we can plan on a carpet to install next month. \EI
\I  General restocking of needed items
\I  Re-organize the janitor closet
    \BI Feel free to throw away what does not appear useful
        (e.g. bottles with only a little solution remaining). \EI
\I  Annex
    \BI Sweep floors
    \I  Clean and restock bathroom (including better solution for hand soap) \EI
\I  General outdoor cleaning
    \BI Leaves in front of main church entryway.
    \I  General weeding in the lawn area between the main building and annex. \EI
\I  Annex soundproofing
    \BI Research creative, cost-efficient ways to add soundproofing to the
        Annex dining room. \EI
\I  Remove or grind down the metal peice protruding from Annex.
\I  Purchase and install rubber ring-boots for vent flashings 
    (pipes sticking out of roof) near the roof by the main entrance 
    of the main building.
    \BI Diameter of pipes (flashings) and approximate degree of roof 
        need to first be measured in order to purchase the correct rubber boots.
    \I  Rubber boots then need to be installed. \EI
\I  Finish trimming hedges around the parking lot. 
\EI
\end{document}