### 2013-09-15 sandbox/pbc-workday/2013-09.tex
iNew workday list.

\documentclass[12pt]{article}
\newcommand{\BI}{\begin{enumerate}\item}
\newcommand{\I}{\item}
\newcommand{\EI}{\end{enumerate}}
\title{Work Day}
\begin{document}
\maketitle
Items are listed in order of \emph{approximate} priority. 
\BI Main Sanctuary
    \BI Clean carpet using a carpet cleaner
    \I  Clean pews using Pastor Josh's carpet/upholstery machine
    \I  Vacuum pews and carpet
    \I  General cleaning (remove old bulletins...) \EI
\I  Bathrooms: Regular cleaning and restocking
\I  Tree trimming
    \BI Trees between sidewalk and parking lot
    \I  Trees near sign at parking lot entrance
    \I  Hedges around the parking lot \EI
\I  Annex
    \BI Sweep floors
    \I  Clean and restock bathroom (including better solution for hand soap) \EI
\I  Replace fluorescent lights in pastor's office
    \BI Bulb type is F50, soft white (may need to purchase at Lowe's) \EI
\I  General outdoor cleaning
    \BI Leaves in front of main church entryway.
    \I  General weeding in the lawn area between the main building and annex. \EI
\I  Annex soundproofing
    \BI Research creative, cost-efficient ways to add soundproofing to the
        Annex dining room. \EI
\I  Remove or grind down the metal peice protruding from Annex.
\I  Purchase and install rubber ring-boots for vent flashings 
    (pipes sticking out of roof) near the roof by the main entrance 
    of the main building.
    \BI Diameter of pipes (flashings) and approximate degree of roof 
        need to first be measured in order to purchase the correct rubber boots.
    \I  Rubber boots then need to be installed. \EI
\EI
\end{document}